---
title: "Further cleaning"
output: html_notebook
---
Some further data cleaning commands to get an input for the machine learning algorithm

Yday function is to covert the date to the doy
```{r}
climate_imputed<-read.table('../data/kyoto_climate_imputed.csv' ,sep=",", header=TRUE)
climate_imputed$DATE = as.Date(climate_imputed$DATE )
climate_imputed$doy = yday(climate_imputed$DATE)
str(climate_imputed)

```

Read the bloom data and combine it with the weather data
Here you can have a look at the data and make sure remove all the redundant information and the NA value to minumize the data size.
This function is used to detect the na is.na part
test <- c(1,2,3,NA) is.na(test)
This function is to use remove the na
ompleterecords <- na.omit(datacollected)

```{r}
bloom_result<-read.table('../data/kyoto.csv' ,sep=",", header=TRUE)
bloom_result$bloom_date = as.Date(bloom_result$bloom_date)
bloom_test=bloom_result %>% right_join(climate_imputed,by="year")
bloom_test<-mutate(bloom_test,Status = case_when(
    DATE>=bloom_date ~ 1,
    DATE<bloom_date ~ 0
    )) 
#bloom_test$Status# = as.numeric(bloom_result$Status)
```